---
title: 'Comparing betas between runwise and subj-wise (DMCC2 batch) GLMs'
author: "michael freund"
date: "15 April 2020"
output:
  html_document:
  toc: true
highlight: zenburn
---
  
```{r setup, include = FALSE}

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = 'center',
  fig.width = 11.5, fig.fullwidth = TRUE
)

set.seed(0)

library(here)
library(magrittr)
library(dplyr)
library(data.table)
library(mikeutils)
library(lme4)
library(ggplot2)
library(ggbeeswarm)
library(grid)
library(gridExtra)
library(cowplot)
library(cifti)
library(gifti)
library(abind)

## settings ----

theme_set(theme_classic(base_size = 8))

## data ----

pretrial <- fread(here("data", "pretrial_behavior.csv"))  ## behavior and events
pretrial.subjsum <- fread(here("data", "pretrial_subjsumm.csv"))  ## summary of behavior and events / subj

## paths ----

sessi <- c("baseline", "proactive")
sessi.short <- c("Bas", "Pro")
glms <- "Congruency_EVENTS_censored"
n.knots <- 6
regressors <- c("PC50InCon", "biasInCon", "PC50Con", "biasCon")
networks <- c("Vis", "DorsAttn", "SalVentAttn", "Cont", "Default", "SomMot", "Limbic")

## atlases ----

parcellation <- read_atlas("schaefer400")

nodename <- Sys.info()["nodename"]
if (nodename == "ccplinux1") {
  dir.atlas <- "/data/nil-external/ccp/freund/atlases"
  dir.schaefer <- "/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/ATLASES/"
} else if (nodename == "CCP-FREUND") {
  dir.atlas <- "C:/local/atlases"
  dir.schaefer <- dir.atlas
}

hcp <- list(
  L  = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
  ),
  R = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")
  )
)

over <- list(
  L = parcellation$atlas[1:(nrow(parcellation$atlas) / 2)], 
  R = parcellation$atlas[(nrow(parcellation$atlas) / 2):nrow(parcellation$atlas)]
)

# schaefer <- list(
#   L = read_gifti2matrix(
#     "/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/ATLASES/Schaefer2018_400Parcels_7Networks_order_L.label.gii"
#     ) %>% c,
#   R = read_gifti2matrix(
#     "/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/ATLASES/Schaefer2018_400Parcels_7Networks_order_R.label.gii"
#     ) %>% c
# )

## stats ----


a <- readRDS(here("out", "qc", paste0("similarity-to-ccphcp-glms_shaefer400_2tr1knot.rds")))

## remove all NA subjs ----

has.stats <- apply(a, "subj", function(.) !any(is.na(c(.))))
a <- a[, , , , has.stats, ]

## wrangle to data.frame ----

d <- reshape2::melt(a, value.name = "simil") %>% rename(session = sess)

d$tr <- as.numeric(gsub("knot", "", d$knot)) * 2  ## knots to TRs

d$network <- ""  ## add networks
for (network.i in networks) d$network[grepl(network.i, d$roi)] <- network.i

d$trial.type <- gsub("PC50|bias", "", d$param)  ## create variables
d$pc <- gsub("InCon|Con", "", d$param)

d$trial.type <- factor(d$trial.type, levels = c("Con", "InCon"))  ## to factors
d$pc <- factor(d$pc, levels = c("PC50", "bias"))
d$session <- factor(d$session, levels = c("baseline", "proactive"))



```

## all data (no aggregation)

```{r}

p.density.corr <- d %>%
  filter(measu == "corr") %>%
  ggplot(aes(x = simil)) +
  geom_density(fill = "firebrick") +
  facet_grid(vars(param), vars(session)) +
  labs(x = "linear correlation")

p.density.eucl <- d %>%
  filter(measu == "eucl") %>%
  ggplot(aes(x = simil)) +
  geom_density(fill = "firebrick") +
  facet_grid(vars(param), vars(session)) +
  labs(x = "euclidean distance")

grid.arrange(p.density.corr, p.density.eucl, ncol = 2, top = "pattern similarity across all parcels, knots, and subjs")

p.boxplot.corr <- d %>%
  filter(measu == "corr") %>%
  ggplot(aes(y = simil)) +
  geom_boxplot(fill = "firebrick", width = 0.25) +
  facet_grid(vars(param), vars(session)) +
  labs(y = "linear correlation")

p.boxplot.eucl <- d %>%
  filter(measu == "eucl") %>%
  ggplot(aes(y = simil)) +
  geom_boxplot(fill = "firebrick", width = 0.25) +
  facet_grid(vars(param), vars(session)) +
  labs(y = "euclidean distance")

grid.arrange(p.boxplot.corr, p.boxplot.eucl, ncol = 2, top = "pattern similarity across all parcels, knots, and subjs")

```

## by knot (no aggregation)

```{r}

p.allparcels.pearson <- d %>%
  filter(measu == "corr") %>%
  ggplot(aes(knot, simil)) +
  geom_violin(width = 0.5, fill = "firebrick") +
  facet_grid(vars(param), vars(session)) +
  labs(title = "linear correlation", x = "TR (post stimulus onset)", y = "beta")

p.allparcels.euclidean <- d %>%
  filter(measu == "eucl") %>%
  ggplot(aes(knot, simil)) +
  geom_violin(width = 0.5, fill = "firebrick") +
  facet_grid(vars(param), vars(session)) +
  labs(title = "euclidean distance", x = "TR (post stimulus onset)", y = "beta")

grid.arrange(p.allparcels.pearson, p.allparcels.euclidean, ncol = 2, top = "all parcels")

```

Dissimilar patterns across all TRs, in all sessions.


## by subject (aggregated across TR)

```{r}

d %>%
  filter(measu == "corr") %>%
  group_by(roi, param, session, subj) %>%
  summarize(simil = tanh(mean(atanh(simil)))) %>%
  ggplot(aes(subj, simil)) +
  geom_boxplot(width = 0.5, fill = "firebrick") +
  facet_grid(vars(param), vars(session)) +
  labs(y = "linear correlation", x = "subj") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 0.5))

d %>%
  filter(measu == "eucl") %>%
  group_by(roi, param, session, subj) %>%
  summarize(simil = mean(simil)) %>%
  ggplot(aes(subj, simil)) +
  geom_boxplot(width = 0.5, fill = "firebrick") +
  facet_grid(vars(param), vars(session)) +
  labs(y = "euclidean distance", x = "subj") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 0.5))


```

Seems to be relatively dissimilar patterns in each subject as well.
Perhaps some variability by subject\*session.


## by parcel (aggregated across TR and subj)

```{r}

d %>%
  filter(measu == "corr") %>%
  group_by(roi, session, subj, network) %>%
  summarize(simil = tanh(mean(atanh(simil)))) %>%
  ggplot(aes(roi, simil)) +
  geom_boxplot(aes(fill = network)) +
  scale_fill_brewer(type = "qual", palette = 2) +
  facet_grid(rows = vars(session), scale = "free") +
  labs(y = "linear correlation", x = "parcel") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 0.5))

```


## similarities of subjs with no errors

```{r}

errors <- pretrial %>%
  group_by(subj, session) %>%
  summarize(error = mean(1 - acc, na.rm = TRUE)) %>%
  mutate(session = ifelse(session == "bas", "baseline", "proactive"))

d %>%
  filter(measu == "corr") %>%
  group_by(subj, session, param) %>%
  summarize(simil = tanh(mean(atanh(simil)))) %>%
  inner_join(errors, by = c("subj", "session")) %>%
  group_by(session, param) %>%
  summarize(r = cor(error, simil, method = "spearman"))

d %>%
  filter(measu == "corr") %>%
  group_by(subj, session, param) %>%
  summarize(simil = tanh(mean(atanh(simil)))) %>%
  inner_join(errors, by = c("subj", "session")) %>%
  ggplot(aes(rank(error), rank(simil))) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(vars(session), vars(param))


subjs.no.errors <- errors %>% 
  filter(error == 0) %>%
  with(table(subj)) %>%
  as.data.frame %>%
  filter(Freq == 2) %>% 
  pull(subj) %>%
  as.character


d %>%
  filter(measu == "corr", subj %in% subjs.no.errors) %>%
  ggplot(aes(y = simil)) +
  geom_boxplot() +
  facet_grid(vars(session), vars(param)) +
  labs(y = "linear correlation")

d %>%
  filter(measu == "corr", subj %in% subjs.no.errors) %>%
  View


d %>%
  filter(measu == "corr", subj %in% subjs.no.errors) %>%
  group_by(session, param, roi, network) %>%
  summarize(simil = tanh(mean(atanh(simil)))) %>%
  ggplot(aes(roi, simil)) +
  geom_point(aes(color = network)) +
  facet_grid(vars(param), vars(session)) +
  labs(y = "linear correlation")



d %>%
  filter(measu == "corr", subj %in% subjs.no.errors) %>%
  group_by(subj, session) %>%
  # group_by(subj, session, param) %>%
  summarize(simil = tanh(mean(atanh(simil)))) %>%
  left_join(errors, by = c("subj", "session")) %>%
  ggplot(aes(rank(error), rank(simil))) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(cols = vars(session))
  # facet_grid(vars(param), vars(session))





```


## simulation


```{r}


nsim <- 1E5

set.seed(0)
n <- 100
X <- cbind(1, rnorm(x))
signal <- x*2 + 10
split1 <- sample.int(n, n/2)
split2 <- -split1

b.full <- matrix(NA, nrow = nsim, ncol = 2)
colnames(b.full) <- c("b0", "b1")
  
b.piecewise <- matrix(NA, nrow = nsim, ncol = 2)
colnames(b.piecewise) <- c("int", "x")
  
for (ii in seq_len(nsim)) {
  
  y <- signal + rnorm(n)

  b.full[ii, ] <- coef(.lm.fit(X, y))
  
  coef.split1 <- coef(.lm.fit(X[split1, ], y[split1]))
  coef.split2 <- coef(.lm.fit(X[split2, ], y[split2]))
  
  b.piecewise[ii, ] <- (coef.split1 + coef.split2) / 2
  
}

  
## means

colMeans(b.full)
colMeans(b.piecewise)

colMeans(b.full) - colMeans(b.piecewise)

## variance

apply(b.full, 2, var)
apply(b.piecewise, 2, var)

apply(b.piecewise, 2, var) - apply(b.full, 2, var)


cor(b.piecewise, b.full)

plot(b.piecewise[, 1], b.full[, 1])
plot(b.piecewise[, 2], b.full[, 2])




```

Piecewise method is unbiased but estimates have larger variance.
Question: is this increase in variance substantial enough to explain the low correlations observed?

## modeling

```{r}

d1 <- d %>% 
  filter(measu == "corr") %>%
  group_by(param, roi, subj, session) %>%
  summarize(simil = mean(atanh(simil)))

fit1 <- lm(simil ~ param + roi + subj + session, d1)

fit.0subj <- update(fit1, . ~ . - subj)
fit.0roi <- update(fit1, . ~ . - roi)
fit.0sess <- update(fit1, . ~ . - session)
fit.0param <- update(fit1, . ~ . - param)


anova(fit1, fit.0subj)  ## variability across subjs
anova(fit1, fit.0roi)  ## across rois
anova(fit1, fit.0sess)  ## no session diffs
anova(fit1, fit.0param)  ## across params


```

