---
title: 'Exploring similarity measures'
author: "michael freund"
date: "3/20/2020"
output:
  html_document:
    toc: true
    highlight: zenburn
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = 'center',
  fig.width = 11.5, fig.fullwidth = TRUE
  )

set.seed(0)

library(here)
library(magrittr)
library(dplyr)
library(data.table)
library(mikeutils)
library(lme4)
library(ggplot2)
# library(ggbeeswarm)
library(grid)
library(gridExtra)
library(cowplot)
library(cifti)
library(gifti)
# library(nifti)

theme_set(theme_classic(base_size = 8))

## variables

atlas <- read_atlas("schaefer400")

nodename <- Sys.info()["nodename"]
if (nodename == "ccplinux1") {
  dir.atlas <- "/data/nil-external/ccp/freund/atlases"
} else if (nodename == "CCP-FREUND") {
  dir.atlas <- "C:/local/atlases"
}

dir.ccp.hcp <- "/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/HCP_SUBJECTS_BACKUPS/AFNI_ANALYSIS"
dir.subsubj <- "/data/nil-external/ccp/witzermanm/AFNI_ANALYSIS_SUBSUBJECT"
dir.results <- file.path(dir.subsubj, "RESULTS_RUNWISE")

pretrial <- fread(here("data", "pretrial_behavior.csv"))
pretrial.subjsum <- fread(here("data", "pretrial_subjsumm.csv"))

subjs.analysis <- unique(pretrial.subjsum[subj.set == "analysis"]$subj)
subjs.development <- unique(pretrial.subjsum[subj.set == "development"]$subj)
subjs.bad <- unique(pretrial.subjsum[subj.set == "bad"]$subj)

## vanilla RSA:

r.vn.corr <- readRDS(here("out", "rsa", "rmatrix_vanilla_corr_shaefer400_baseline_Congruency_EVENTS_censored.rds"))
r.vn.eucl <- readRDS(here("out", "rsa", "rmatrix_vanilla_eucl_shaefer400_baseline_Congruency_EVENTS_censored.rds"))
r.vn.neuc <- readRDS(here("out", "rsa", "rmatrix_vanilla_neuc_shaefer400_baseline_Congruency_EVENTS_censored.rds"))

subjs <- dimnames(r.vn.corr)$subj
subjs.analysis <- intersect(subjs.analysis, subjs)
subjs.development <- intersect(subjs.development, subjs)
subjs.bad <- intersect(subjs.bad, subjs)

# r.vn <- r.vn[, , subjs.development, , , , , , ]  ## HOLD OUT DATA!!!
r.vn.corr <- r.vn.corr[, , , , , subjs.analysis]
r.vn.eucl <- r.vn.eucl[, , , , , subjs.analysis]
r.vn.neuc <- r.vn.neuc[, , , , , subjs.analysis]

parcellation <- read_atlas("schaefer400")
hcp <- list(
  L  = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
    ),
  R = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")
    )
)
over <- list(
  L = parcellation$atlas[1:(nrow(parcellation$atlas) / 2)], 
  R = parcellation$atlas[(nrow(parcellation$atlas) / 2):nrow(parcellation$atlas)]
  )

schaefer <- list(
  L = read_gifti2matrix(
    "/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/ATLASES/Schaefer2018_400Parcels_7Networks_order_L.label.gii"
    ) %>% c,
  R = read_gifti2matrix(
    "/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/ATLASES/Schaefer2018_400Parcels_7Networks_order_R.label.gii"
    ) %>% c
)

# list.files("/data/nil-bluearc/ccp-hcp/DMCC_ALL_BACKUPS/ATLASES")

```

## About

Goal of this analysis is to compare measures of similarity and denoising procedures in their performance.

Performance criteria for include sensitivity (anatomical coverage), effect sizes (RSA-model fit), and pattern reliability (cross-run correlation).

To mitigate overfitting, these analyses are conducted in a set of subjects that will be held-out from the "hypothesis-testing" set used for the main analyses.

Similarity measures considered:

* Pearson's correlation coefficient
* Euclidean distance (negated)

Denoising procedures considered:

* "Raw", i.e., untransformed beta estimates
* Spatially pre-whitened beta estimates
  * using residual covariance matrix
  * shrunken towards diagonal via Ledoit-Wolfe (sp?) procedure
  * averaged across runs for single whitening matrix
  * sqrt-ed and applied to each beta vector separately

## stats

```{r stats}

## define models ----

dnames <- dimnames(r.vn.corr)$.row
is.incon <- grepl("InCon", dnames)
is.congr <- grepl("PC50Con|biasCon", dnames)
is.pc50  <- grepl("PC50", dnames)
is.bias  <- grepl("bias", dnames)
is.run1  <- grepl("run1", dnames)
is.run2  <- grepl("run2", dnames)


m <- diag(length(dnames))
dimnames(m) <- dimnames(r.vn.corr)[c(".row", ".col")]
m[] <- 0
lt <- lower.tri(m)

m.incon <- m
m.incon[is.incon, is.incon] <- 1
m.incon[is.congr, is.congr] <- 0
m.incon[is.congr, is.incon] <- -1
m.incon[is.incon, is.congr] <- -1
m.incon[is.run1, is.run1] <- 0
m.incon[is.run2, is.run2] <- 0

qcor(m.incon, .title = "cross-run incongruency model")

v.incon <- m.incon[lt]
v.incon[v.incon == 1] <- 1 / sum(v.incon == 1)
v.incon[v.incon == -1] <- -1 / sum(v.incon == -1)


## get subject-level stats (RSA-model fits) ---

r.vn.corr <- atanh(r.vn.corr)
r.vn.eucl <- -r.vn.eucl
r.vn.neuc <- -r.vn.neuc

stats.subjs.corr <- apply(
  r.vn.corr, 
  c("subj", "roi", "norma", "knot"),
  function(x) x[lt] %*% v.incon
  )
stats.subjs.eucl <- apply(
  r.vn.eucl, 
  c("subj", "roi", "norma", "knot"),
  function(x) x[lt] %*% v.incon
  )
stats.subjs.neuc <- apply(
  r.vn.neuc, 
  c("subj", "roi", "norma", "knot"),
  function(x) x[lt] %*% v.incon
  )

stats.subjs <- rbind(
  reshape2::melt(stats.subjs.corr) %>% mutate(measu = "corr"),
  reshape2::melt(stats.subjs.eucl) %>% mutate(measu = "eucl"),
  reshape2::melt(stats.subjs.neuc) %>% mutate(measu = "neuc")
)
stats.subjs$num.roi <- match(stats.subjs$roi, atlas$key)
stats.subjs %<>% mutate(hemi = substr(roi, 1, 1))

## remove subjs with no glm:

subjs.nofit <- stats.subjs %>%
  filter(is.na(value)) %>%
  pull(subj) %>%
  as.character %>%
  unique

stats.subjs %<>% filter(!subj %in% subjs.nofit)

## get group-level stats ----

stats.group <- stats.subjs %>%
  filter(knot %in% c("knot2", "knot3")) %>%
  group_by(subj, num.roi, roi,hemi, measu, norma) %>%
  summarize(value = mean(value)) %>%
  group_by(num.roi, roi, hemi, measu, norma) %>%
  summarize(
    p = wilcox.test(value, alternative = "greater")$p.value,
    w = wilcox.test(value, alternative = "greater")$statistic,
    m = mean(value)
  ) %>%
  group_by(measu, norma) %>%
  mutate(p.fdr = p.adjust(p, method = "fdr"))

```



## coverage

* displayed on brains is RSA-model fit statistics (mean contrast)
* average of knots 2 and 3
* order: correlation, euclidean, normalized euclidean; raw first, then prewhitened
* for thresholded brains: alpha = 0.05 FDR corrected across all 400 parcels

#### unthresholded

```{r coverage_unthresh}

stats.group %>%
  filter(measu == "corr", norma == "raw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  filter(measu == "eucl", norma == "raw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  filter(measu == "neuc", norma == "raw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  filter(measu == "corr", norma == "prw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  filter(measu == "eucl", norma == "prw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  filter(measu == "neuc", norma == "prw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)


```


#### thresholded

```{r coverage_thresh}

stats.group %>%
  mutate(m = ifelse(p.fdr < 0.05, m, 0)) %>%
  filter(measu == "corr", norma == "raw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  mutate(m = ifelse(p.fdr < 0.05, m, 0)) %>%
  filter(measu == "eucl", norma == "raw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)
 
stats.group %>%
  mutate(m = ifelse(p.fdr < 0.05, m, 0)) %>%
  filter(measu == "neuc", norma == "raw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  mutate(m = ifelse(p.fdr < 0.05, m, 0)) %>%
  filter(measu == "corr", norma == "prw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

stats.group %>%
  mutate(m = ifelse(p.fdr < 0.05, m, 0)) %>%
  filter(measu == "eucl", norma == "prw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)
 
stats.group %>%
  mutate(m = ifelse(p.fdr < 0.05, m, 0)) %>%
  filter(measu == "neuc", norma == "prw") %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)
 
```


#### comparison of test statistics

```{r test_stats}

stats.group.wide <- stats.group %>%
  select(roi, measu, norma, w, p.fdr, m) %>%
  as.data.table %>%
  dcast(roi + measu ~ norma, value.var = c("w", "p.fdr", "m")) %>%
  mutate(
    is.sig.raw = p.fdr_raw < 0.05,
    is.sig.prw = p.fdr_prw < 0.05,
    sig = ifelse(is.sig.raw & is.sig.prw, "both", ifelse(is.sig.raw, "raw", ifelse(is.sig.prw, "prw", "none")))
  )

grid.arrange(
  
  stats.group.wide %>%
    filter(measu == "corr") %>%
    ggplot(aes(w_raw, w_prw)) +
    geom_abline() +
    geom_point(aes(color = sig)) +
    labs(title = "correlation") +
    coord_cartesian(xlim = c(0, 750), ylim = c(0, 750)),
  
  stats.group.wide %>%
    filter(measu == "eucl") %>%
    ggplot(aes(w_raw, w_prw)) +
    geom_abline() +
    geom_point(aes(color = sig)) +
    labs(title = "euclidean") +
    coord_cartesian(xlim = c(0, 750), ylim = c(0, 750)),
  
  stats.group.wide %>%
    filter(measu == "neuc") %>%
    ggplot(aes(w_raw, w_prw)) +
    geom_abline() +
    geom_point(aes(color = sig)) +
    labs(title = "normalized euclidean") +
    coord_cartesian(xlim = c(0, 750), ylim = c(0, 750)),
  
  ncol = 3
  
)

```

