---
title: 'Defining ROIs: vanilla RSA'
author: "mike freund"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    highlight: zenburn

---

```{r setup, include = FALSE}

measure.i <- "neuc"

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = 'center',
  fig.width = 11.5, fig.fullwidth = TRUE,
  cache.path = paste0(measure.i, "/cache/"),
  fig.path = paste0(measure.i, "/figs/")
)

set.seed(0)

library(here)
library(magrittr)
library(dplyr)
library(tidyr)
library(data.table)
library(mikeutils)
library(lme4)
library(lmerTest)
library(ggplot2)
library(ggbeeswarm)
library(ggridges)
library(viridis)
library(grid)
library(gridExtra)
library(cowplot)
library(cifti)
library(gifti)
library(abind)

source(here("src", "setup.R"))

## settings ----

theme_set(theme_minimal(base_size = 14))

## functions ----

plot_matrix <- function(d) {
  
  if (is.matrix(d)) d <- reshape2::melt(d)
  
  d %>%
    mutate(.col = factor(.col, levels = rev(unique(.col)))) %>%
    ggplot(aes(.row, .col, fill = value)) +
    geom_tile() +
    # scale_fill_viridis() +
    scale_fill_gradient(low = "black", high = "white") +
    theme(
      # legend.position = "none", 
      panel.grid = element_blank(), 
      # panel.background = element_blank(),
      axis.text.x = element_blank(),
      # axis.ticks = element_blank(), 
      axis.title = element_blank()
      )

}

tidy_lmer <- function(x) {
  coefs <- as.data.frame(coef(summary(x)))
  coefs$term <- rownames(coefs)
  dplyr::rename(coefs, estimate = "Estimate", se = "Std. Error", "statistic" = "t value", p.value = "Pr(>|t|)")
}

## data ----

## parcel-wise similarity matrices

vanil.2tr1knot <- abind(
  readRDS(
    here(
      "out", "rsa", "observed", 
      paste0("rmatrix_vanilla_", measure.i, "_shaefer400_baseline_Congruency_EVENTS_censored.rds")
      )
    ),
  readRDS(
    here(
      "out", "rsa", "observed", 
      paste0("rmatrix_vanilla_", measure.i, "_shaefer400_proactive_Congruency_EVENTS_censored.rds")
      )
    ),
  rev.along = 0
)

names(dimnames(vanil.2tr1knot)) <- c(".row", ".col", "norma", "knot", "parcel", "subj", "session")
dimnames(vanil.2tr1knot)$session <- c("baseline", "proactive")

vanil.1tr1knot <- abind(
  readRDS(
    here(
      "out", "rsa", "observed", 
      paste0("rmatrix_vanilla_", measure.i, "_shaefer400_baseline_Congruency_EVENTS_tentzero01210_censored.rds")
      )
    ),
  readRDS(
    here(
      "out", "rsa", "observed",
      paste0("rmatrix_vanilla_", measure.i, "_shaefer400_proactive_Congruency_EVENTS_tentzero01210_censored.rds"))
    ),
  rev.along = 0
)

names(dimnames(vanil.1tr1knot)) <- c(".row", ".col", "norma", "knot", "parcel", "subj", "session")
dimnames(vanil.1tr1knot)$session <- c("baseline", "proactive")

## filter subjs by those with data

has.stats.2tr1knot <- apply(vanil.2tr1knot, "subj", function(.) !any(is.na(c(.))))
has.stats.1tr1knot <- apply(vanil.1tr1knot, "subj", function(.) !any(is.na(c(.))))
has.stats <- has.stats.2tr1knot & has.stats.1tr1knot
subjs.with.stats <- names(has.stats)[has.stats]

## TODO
## these subjects do not have stats
# setdiff(c(subjs.analysis, subjs.development), subjs.with.stats)
## could be due to
##  - no 3d+t gifti -> have RAs run
##  - no beta gifti -> diagnose problem and fit

subjs.analysis <- intersect(subjs.analysis, subjs.with.stats)
subjs.development <- intersect(subjs.development, subjs.with.stats)
subjs.bad <- intersect(subjs.bad, subjs.with.stats)


## subset data ----

vanil.2tr1knot <- vanil.2tr1knot[, , , , , subjs.analysis, ]
vanil.1tr1knot <- vanil.1tr1knot[, , , , , subjs.analysis, ]

## atlases ----

hcp <- list(
  L  = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
  ),
  R = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")
  )
)

over <- list(
  L = parcellation$atlas[1:(nrow(parcellation$atlas) / 2)], 
  R = parcellation$atlas[(nrow(parcellation$atlas) / 2):nrow(parcellation$atlas)]
)

schaefer <- list(
  L = read_gifti2matrix(
    file.path(dir.schaefer, "Schaefer2018_400Parcels_7Networks_order_L.label.gii")
    ) %>% c,
  R = read_gifti2matrix(
    file.path(dir.schaefer, "Schaefer2018_400Parcels_7Networks_order_R.label.gii")
    ) %>% c
)

## palettes ----

colors.glm <- c("2tr1knot" = "#d95f02", "1tr1knot" = "#1b9e77")
colors.celltype <- c(I_I = "#e41a1c", C_C = "#377eb8", I_C = "#4daf4a")

## wrangle ----


## transformations

if (measure.i == "corr") {
  
  ## fisher transform correlations
  
  vanil.1tr1knot <- atanh(vanil.1tr1knot)  ## r to z
  vanil.2tr1knot <- atanh(vanil.2tr1knot)

} else if (measure.i %in% c("eucl", "neuc")) {

  ## negate (for comparison to correlation)

  vanil.1tr1knot <- -vanil.1tr1knot
  vanil.2tr1knot <- -vanil.2tr1knot
  
}

inds.pc50.run1 <- grep("PC50.*_run1", conds.run)
inds.pc50.run2 <- grep("PC50.*_run2", conds.run)
vanil.1tr1knot.pc50 <- vanil.1tr1knot[inds.pc50.run2, inds.pc50.run1, , , , , ]
vanil.2tr1knot.pc50 <- vanil.2tr1knot[inds.pc50.run2, inds.pc50.run1, , , , , ]

## unwrap to lower triangle vectors

vanil.1tr1knot.pc50.d <- reshape2::melt(vanil.1tr1knot.pc50, value.name = "simil", na.rm = TRUE) %>% 
  mutate(glm = "1tr1knot")
vanil.2tr1knot.pc50.d <- reshape2::melt(vanil.2tr1knot.pc50, value.name = "simil", na.rm = TRUE) %>% 
  mutate(glm = "2tr1knot")

## bind and add factor cols

vanil.2tr1knot.pc50.d$tr <- as.numeric(gsub("knot", "", vanil.2tr1knot.pc50.d$knot)) * 2  ## knots to TRs
vanil.1tr1knot.pc50.d$tr <- as.numeric(gsub("knot", "", vanil.1tr1knot.pc50.d$knot))
vanil.pc50 <- bind_rows(vanil.1tr1knot.pc50.d, vanil.2tr1knot.pc50.d)  ## bind different glms

vanil.pc50$network <- ""  ## add networks
for (network.i in networks) vanil.pc50$network[grepl(network.i, vanil.pc50$parcel)] <- network.i

vanil.pc50$celltype <- ifelse(
  grepl("PC50InCon", vanil.pc50$.row) & grepl("PC50InCon", vanil.pc50$.col), "I_I",
  ifelse(
    grepl("PC50Con", vanil.pc50$.row) & grepl("PC50Con", vanil.pc50$.col), "C_C", "I_C"
  )
)

vanil.pc50 %<>% mutate(parcel.num = match(parcel, parcellation$key))

## take out trash

rm(
  vanil.1tr1knot.pc50, vanil.1tr1knot.pc50.d,
  vanil.2tr1knot.pc50, vanil.2tr1knot.pc50.d
  )

gc()

vanil.pc50 %<>% filter(norma == "raw") %>% select(-norma)  ## raw values only (not prewhitened)


## stats ----

vanil.pc50.parcel.contrast <- vanil.pc50 %>%
  
  group_by(session, network, tr, subj, parcel, celltype, glm, parcel.num) %>%
  summarize(simil = mean(simil)) %>%  ## average within celltype
  
  pivot_wider(names_from = "celltype", values_from = "simil") %>%  ## get contrast
  mutate(simil.contrast = I_I - I_C)

vanil.pc50.dmcc34 <- vanil.pc50 %>% filter(parcel.num %in% dmcc34)

```


# about

1. A cursory vanilla RSA on baseline and proactive data, PC50 items.
2. Identify a set of ROIs via RSA, separately in baseline and proactive, that respond to incongruency status.

* __subjs__:

* __similarity measures__:

* __glms__:

* __sessions__:

* __conditons / trial types__:



# data at a glance


```{r distributions, fig.height = 7}

## https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html

grid.arrange(
  
  vanil.pc50 %>%
    
    ggplot(aes(simil, y = network, fill = network, height = stat(density))) +
    geom_density_ridges(stat = "density", alpha = 0.8, rel_min_height = 0.01) +
    
    facet_grid(vars(session, glm)) +
    scale_color_brewer(type = "qual") +
    theme_minimal() +
    theme(axis.text.y = element_text(vjust = 0), legend.position = "none"),
  
  vanil.pc50 %>%
    
    ggplot(aes(network, simil, fill = network)) +
    geom_boxplot(width = 0.25, outlier.alpha = 0.5, outlier.size = 0.5) +
    
    facet_grid(vars(session, glm)) +
    scale_color_brewer(type = "qual") +
    theme_minimal() +
    theme(legend.position = "none"),
  
  ncol = 2
  
)

```

__Mean (whole-brain) similarity matrices (2tr1knot only):__

```{r matrices, fig.height = 10}

allsim.2tr1knot.bas <- tanh(apply(atanh(vanil.2tr1knot[, , "raw", , , , "baseline"]), c(".row", ".col"), mean))
allsim.2tr1knot.pro <- tanh(apply(atanh(vanil.2tr1knot[, , "raw", , , , "proactive"]), c(".row", ".col"), mean))

grid.arrange(
  
  plot_matrix(allsim.2tr1knot.bas) + labs(title = "baseline"),
  plot_matrix(allsim.2tr1knot.pro) + labs(title = "proactive"),
  
  plot_matrix(allsim.2tr1knot.bas[1:4, 5:8]) + labs(title = "baseline, between-run"),
  plot_matrix(allsim.2tr1knot.pro[1:4, 5:8]) + labs(title = "proactive, between-run"),
  
  plot_matrix(allsim.2tr1knot.bas[inds.pc50.run1, inds.pc50.run2]) + 
    labs(title = "baseline, between-run PC50") + theme(axis.text.x = element_text()),
  plot_matrix(allsim.2tr1knot.pro[inds.pc50.run1, inds.pc50.run2]) + 
    labs(title = "proactive, between-run PC50") + theme(axis.text.x = element_text()),

  ncol = 2
  
)

```

__Here I will focus only on the BETWEEN-RUN PC50 cells.__


# event-related curves

## all parcels

### by network

__aggregated across parcels within network__

```{r curves_networ, fig.height = 12}

## z values, by network

vanil.pc50 %>%
  
  group_by(session, network, tr, subj, celltype, glm) %>%
  summarize(simil = mean(simil)) %>%  ## average across parcels
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot", geom = "ribbon", alpha = 0.25) +
  stat_summary(fun.y = "mean", geom = "line", size = 1) +
  
  facet_grid(vars(network), vars(session, glm)) +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "pairwise similarities over target epoch", 
    y = "similarity", 
    caption = "bands: bootstrapped 95% CI of across-subject mean"
    )

## contrast, by network

vanil.pc50 %>%
  
  group_by(session, network, tr, subj, celltype, glm) %>%
  summarize(simil = mean(simil)) %>%  ## average across parcels
  
  pivot_wider(names_from = "celltype", values_from = "simil") %>%  ## get contrast
  mutate(simil.contrast = I_I - I_C) %>%
  
  
  ggplot(aes(tr, simil.contrast, color = glm, fill = glm)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot", geom = "ribbon", alpha = 0.25) +
  stat_summary(fun.y = "mean", geom = "line", size = 1) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_fill_manual(values = colors.glm) +
  scale_color_manual(values = colors.glm) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "similarity contrast over target epoch", 
    y = "I-I versus I-C similarity", 
    caption = "bands: bootstrapped 95% CI of across-subject mean"
    )

## t stat, by network

vanil.pc50 %>%
  
  group_by(session, network, tr, celltype, glm, subj) %>%
  summarize(simil = mean(simil)) %>%  ## average across parcels
  
  pivot_wider(names_from = "celltype", values_from = "simil") %>%  ## get contrast
  mutate(simil.contrast = I_I - I_C) %>%
  
  summarize(t.stat = t.test(simil.contrast, alternative = "greater")$statistic) %>%
  
  
  ggplot(aes(tr, t.stat, color = glm)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  geom_line(size = 2) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_color_manual(values = colors.glm) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "none") +
  labs(
    title = "similarity contrast over target epoch", 
    y = "t statistic",
    caption = "dashed line at t = 2"
  )


```


### by parcel

__aggregated across subjects__

#### 2tr1knot

```{r curves_parcel_2tr1knot, fig.height = 12}

## z values, by parcel

vanil.pc50 %>%
  
  filter(glm == "2tr1knot") %>%
  
  group_by(session, network, tr, celltype, parcel) %>%
  summarize(simil = mean(simil)) %>%  ## average across subjects
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.5) +
  
  facet_grid(vars(network), vars(session, celltype)) +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "pairwise similarities over target epoch per parcel | 2tr1knot", 
    y = "similarity"
    )


## contrast, by parcel

vanil.pc50.parcel.contrast %>% 
  
  filter(glm == "2tr1knot") %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(simil.contrast = mean(simil.contrast))  %>% ## average across subjects
  
  ggplot(aes(tr, simil.contrast)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.25) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "similarity contrast over target epoch per parcel | 2tr1knot", 
    y = "I-I versus I-C similarity"
    )

## t stat, by network

vanil.pc50.parcel.contrast %>%
  
  filter(glm == "2tr1knot") %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(t.stat = t.test(simil.contrast, alternative = "greater")$statistic) %>%
  
  group_by(session, network, parcel) %>%
  mutate(is.sig = ifelse(any(t.stat > 3), TRUE, FALSE)) %>%
  
  ggplot(aes(tr, t.stat)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  geom_line(aes(group = parcel, alpha = is.sig), size = 1) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_alpha_manual(values = c("TRUE" = 0.9, "FALSE" = 0.2)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "none") +
  labs(
    title = "similarity contrast over target epoch per parcel | 2tr1knot", 
    y = "t statistic",
    caption = "dark lines are parcels with any TR with t > 3"
  )


```

#### 1tr1knot

```{r curves_parcel_1tr1knot, fig.height = 12}

## z values, by parcel

vanil.pc50 %>%
  
  filter(glm == "1tr1knot") %>%
  
  group_by(session, network, tr, celltype, parcel) %>%
  summarize(simil = mean(simil)) %>%  ## average across subjects
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.5) +
  
  facet_grid(vars(network), vars(session, celltype)) +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 4)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "pairwise similarities over target epoch per parcel | 1tr1knot", 
    y = "similarity"
    )


## contrast, by parcel

vanil.pc50.parcel.contrast %>% 
  
  filter(glm == "1tr1knot") %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(simil.contrast = mean(simil.contrast))  %>% ## average across subjects
  
  ggplot(aes(tr, simil.contrast)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.25) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "similarity contrast over target epoch per parcel | 1tr1knot", 
    y = "I-I versus I-C similarity"
    )

## t stat, by network

vanil.pc50.parcel.contrast %>%
  
  filter(glm == "1tr1knot") %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(t.stat = t.test(simil.contrast, alternative = "greater")$statistic) %>%
  
  group_by(session, network, parcel) %>%
  mutate(is.sig = ifelse(any(t.stat > 3), TRUE, FALSE)) %>%
  
  ggplot(aes(tr, t.stat)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  geom_line(aes(group = parcel, alpha = is.sig), size = 1) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_alpha_manual(values = c("TRUE" = 0.9, "FALSE" = 0.2)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "none") +
  labs(
    title = "similarity contrast over target epoch per parcel | 1tr1knot", 
    y = "t statistic",
    caption = "dark lines are parcels with any TR with t > 3"
  )


```


## DMCC-34

Now just for the DMCC34 parcels.

### by network

__aggregated across parcels within network__

```{r curves_network_dmcc34, fig.height = 12}

## z values, by network

vanil.pc50.dmcc34 %>%
  
  group_by(session, network, tr, subj, celltype, glm) %>%
  summarize(simil = mean(simil)) %>%  ## average across parcels
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot", geom = "ribbon", alpha = 0.25) +
  stat_summary(fun.y = "mean", geom = "line", size = 1) +
  
  facet_grid(vars(network), vars(session, glm)) +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "pairwise similarities over target epoch | dmcc34", 
    y = "similarity", 
    caption = "bands: bootstrapped 95% CI of across-subject mean"
    )

## contrast, by network

vanil.pc50.dmcc34 %>%
  
  group_by(session, network, tr, subj, celltype, glm) %>%
  summarize(simil = mean(simil)) %>%  ## average across parcels
  
  pivot_wider(names_from = "celltype", values_from = "simil") %>%  ## get contrast
  mutate(simil.contrast = I_I - I_C) %>%
  
  
  ggplot(aes(tr, simil.contrast, color = glm, fill = glm)) +
  geom_hline(yintercept = 0) +
  stat_summary(fun.data = "mean_cl_boot", geom = "ribbon", alpha = 0.25) +
  stat_summary(fun.y = "mean", geom = "line", size = 1) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_fill_manual(values = colors.glm) +
  scale_color_manual(values = colors.glm) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "similarity contrast over target epoch | dmcc34", 
    y = "I-I versus I-C similarity", 
    caption = "bands: bootstrapped 95% CI of across-subject mean"
    )

## t stat, by network

vanil.pc50.dmcc34 %>%
  
  group_by(session, network, tr, celltype, glm, subj) %>%
  summarize(simil = mean(simil)) %>%  ## average across parcels
  
  pivot_wider(names_from = "celltype", values_from = "simil") %>%  ## get contrast
  mutate(simil.contrast = I_I - I_C) %>%
  
  summarize(t.stat = t.test(simil.contrast, alternative = "greater")$statistic) %>%
  
  
  ggplot(aes(tr, t.stat, color = glm)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  geom_line(size = 2) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_color_manual(values = colors.glm) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "none") +
  labs(
    title = "similarity contrast over target epoch | dmcc34", 
    y = "t statistic",
    caption = "dashed line at t = 2"
  )


```

### by parcel

__aggregated across subjects__

#### 2tr1knot

```{r curves_parcel_dmcc34_2tr1knot, fig.height = 12}

## z values, by parcel

vanil.pc50.dmcc34 %>%
  
  filter(glm == "2tr1knot") %>%
  
  group_by(session, network, tr, celltype, parcel) %>%
  summarize(simil = mean(simil)) %>%  ## average across subjects
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.5) +
  
  facet_grid(vars(network), vars(session, celltype)) +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "pairwise similarities over target epoch per parcel | 2tr1knot, dmcc34", 
    y = "similarity"
    )


## contrast, by parcel

vanil.pc50.parcel.contrast %>% 
  
  filter(glm == "2tr1knot", parcel.num %in% dmcc34) %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(simil.contrast = mean(simil.contrast))  %>% ## average across subjects
  
  ggplot(aes(tr, simil.contrast)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.25) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "similarity contrast over target epoch per parcel | 2tr1knot, dmcc34", 
    y = "I-I versus I-C similarity"
    )

## t stat, by network

vanil.pc50.parcel.contrast %>%
  
  filter(glm == "2tr1knot", parcel.num %in% dmcc34) %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(t.stat = t.test(simil.contrast, alternative = "greater")$statistic) %>%
  
  group_by(session, network, parcel) %>%
  mutate(is.sig = ifelse(any(t.stat > 3), TRUE, FALSE)) %>%
  
  ggplot(aes(tr, t.stat)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  geom_line(aes(group = parcel, alpha = is.sig), size = 1) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_alpha_manual(values = c("TRUE" = 0.9, "FALSE" = 0.2)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "none") +
  labs(
    title = "similarity contrast over target epoch per parcel | 2tr1knot, dmcc34", 
    y = "t statistic",
    caption = "dark lines are parcels with any TR with t > 3"
  )


```

#### 1tr1knot

```{r curves_parcel_dmcc34_1tr1knot, fig.height = 12}

## z values, by parcel

vanil.pc50.dmcc34 %>%
  
  filter(glm == "1tr1knot") %>%
  
  group_by(session, network, tr, celltype, parcel) %>%
  summarize(simil = mean(simil)) %>%  ## average across subjects
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.5) +
  
  facet_grid(vars(network), vars(session, celltype)) +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "pairwise similarities over target epoch per parcel | 1tr1knot, dmcc34", 
    y = "similarity"
    )


## contrast, by parcel

vanil.pc50.parcel.contrast %>% 
  
  filter(glm == "1tr1knot", parcel.num %in% dmcc34) %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(simil.contrast = mean(simil.contrast))  %>% ## average across subjects
  
  ggplot(aes(tr, simil.contrast)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = parcel), size = 1, alpha = 0.25) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom") +
  labs(
    title = "similarity contrast over target epoch per parcel | 1tr1knot, dmcc34", 
    y = "I-I versus I-C similarity"
    )

## t stat, by network

vanil.pc50.parcel.contrast %>%
  
  filter(glm == "1tr1knot", parcel.num %in% dmcc34) %>%
  
  group_by(session, network, tr, parcel) %>%
  summarize(t.stat = t.test(simil.contrast, alternative = "greater")$statistic) %>%
  
  group_by(session, network, parcel) %>%
  mutate(is.sig = ifelse(any(t.stat > 3), TRUE, FALSE)) %>%
  
  ggplot(aes(tr, t.stat)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  geom_line(aes(group = parcel, alpha = is.sig), size = 1) +
  
  facet_grid(vars(network), vars(session)) +
  
  scale_alpha_manual(values = c("TRUE" = 0.9, "FALSE" = 0.2)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "none") +
  labs(
    title = "similarity contrast over target epoch per parcel | 1tr1knot, dmcc34", 
    y = "t statistic",
    caption = "dark lines are parcels with any TR with t > 3"
  )


```

<!-- ### subject curves within parcel -->

<!-- A closer look at similarity timecourses within these parcels. -->


```{r curves_withinparcel_dmcc34_2tr1knot_allconds, fig.height = 45, eval = FALSE}

## z values, by parcel

vanil.pc50.dmcc34 %>%
  
  filter(glm == "2tr1knot") %>%
  
  group_by(session, network, tr, parcel, subj, celltype) %>%  ## aggregate by celltype
  summarize(simil = mean(simil)) %>%
  
  ggplot(aes(tr, simil, fill = celltype, color = celltype)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = subj), size = 0.5, alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot", geom = "ribbon", alpha = 0.5) +
  stat_summary(fun.y = "mean", geom = "line", size = 1) +
  
  facet_grid(vars(parcel), vars(session, celltype), scales = "free_y") +
  
  scale_fill_manual(values = colors.celltype) +
  scale_color_manual(values = colors.celltype) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom", strip.text.y = element_text(angle = 0, size = 6, hjust = 0, vjust = 0.5)) +
  labs(
    title = "pairwise similarities over target epoch | dmcc34", 
    y = "similarity", 
    caption = "bands: bootstrapped 95% CI of across-subject mean"
    )


```


```{r curves_withinparcel_dmcc34_2tr1knot_contrast, fig.height = 45, fig.width = 7, eval = FALSE}

vanil.pc50.parcel.contrast %>% 
  
  filter(glm == "2tr1knot", parcel.num %in% dmcc34) %>%
  
  ggplot(aes(tr, simil.contrast)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = subj), size = 0.5, alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot", geom = "ribbon", alpha = 0.5) +
  stat_summary(fun.y = "mean", geom = "line", size = 1) +
  
  facet_grid(vars(parcel), vars(session), scales = "free_y") +
  
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  
  theme(legend.position = "bottom", strip.text.y = element_text(angle = 0, size = 6, hjust = 0, vjust = 0.5)) +
  labs(
    title = "similarity contrast over target epoch per parcel | 1tr1knot, dmcc34", 
    y = "similarity", 
    caption = "bands: bootstrapped 95% CI of across-subject mean"
    )


```


# ROI analysis

## models and stats


```{r roi_analysis, fig.height = 15, fig.width = 8}

## fit

vanil.roi <- vanil.pc50.parcel.contrast %>%
  filter(tr > 1, tr < 7) %>%
  split(paste(.$parcel, .$session, .$network, .$glm, sep = "."))

mods.vanil.roi <- lapply(vanil.roi, function(x) lmer(simil.contrast ~ 1 + (1 | subj), x))

## extract

coef.vanil.roi <- lapply(mods.vanil.roi, tidy_lmer)

## format

coef.vanil.roi %<>%
  bind_rows(.id = "model.id") %>%
  cbind(reshape2::colsplit(.$model.id, "\\.", c("parcel", "session", "network", "glm")))

## adjust

coef.vanil.roi %<>%
  group_by(session, glm) %>%
  mutate(p.fdr = p.adjust(p.value, "fdr"))

## add cols for plotting

coef.vanil.roi %<>%
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
    )

## plot

coef.vanil.roi %>%

  select(-p.value, -p.fdr, -estimate, -se, -df, -model.id) %>%
  pivot_wider(names_from = "glm", values_from = statistic, names_prefix = "glm.") %>%

  ggplot(aes(glm.1tr1knot, glm.2tr1knot)) +
  geom_point(
    aes(),
    size = 1.5, shape = 21, fill = "black", color = "grey"
  ) +

  facet_grid(vars(network), vars(session)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline() +

  labs(title = "stroop t statistics, estimated over TRs 2-6")

## list

roi.bas.1tr1knot <- coef.vanil.roi %>%
  filter(
    p.fdr < 0.05, session == "baseline", statistic > 0,
    glm == "1tr1knot"
    ) %>%
  pull(parcel) %>%
  unique

roi.bas.2tr1knot <- coef.vanil.roi %>%
  filter(
    p.fdr < 0.05, session == "baseline", statistic > 0,
    glm == "2tr1knot"
    ) %>%
  pull(parcel) %>%
  unique

roi.pro.1tr1knot <- coef.vanil.roi %>%
  filter(
    p.fdr < 0.05, session == "proactive", statistic > 0,
    glm == "1tr1knot"
    ) %>%
  pull(parcel) %>%
  unique

roi.pro.2tr1knot <- coef.vanil.roi %>%
  filter(
    p.fdr < 0.05, session == "proactive", statistic > 0,
    glm == "2tr1knot"
    ) %>%
  pull(parcel) %>%
  unique



```


At FDR-adjusted $\alpha$ = 0.05


* `r length(roi.bas.2tr1knot)` parcels for 2tr1knot, baseline.

`r roi.bas.2tr1knot`

* `r length(roi.bas.1tr1knot)` parcels for 1tr1knot, baseline.

`r roi.bas.1tr1knot`

* `r length(roi.pro.2tr1knot)` parcels for 2tr1knot, proactive.

`r roi.pro.2tr1knot`

* `r length(roi.pro.1tr1knot)` parcels for 1tr1knot, proactive.

`r roi.pro.1tr1knot`


## brains


* the t statistics from above are plotted on brains here.

## unthresholded

### baseline 1tr1knot

```{r unthresh_baseline_1tr1knot, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "baseline", glm == "1tr1knot") %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

### baseline 2tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "baseline", glm == "2tr1knot") %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

### proactive 1tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "proactive", glm == "1tr1knot") %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

### proactive 2tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "proactive", glm == "2tr1knot") %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```


## thresholded


### baseline 1tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "baseline", glm == "1tr1knot") %>%
  mutate(statistic = ifelse(p.fdr < 0.05 & statistic > 0, statistic, 0)) %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

### baseline 2tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "baseline", glm == "2tr1knot") %>%
  mutate(statistic = ifelse(p.fdr < 0.05 & statistic > 0, statistic, 0)) %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

### proactive 1tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "proactive", glm == "1tr1knot") %>%
  mutate(statistic = ifelse(p.fdr < 0.05 & statistic > 0, statistic, 0)) %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

### proactive 2tr1knot

```{r, fig.height = 7, fig.width = 11}

coef.vanil.roi %>%
  filter(session == "proactive", glm == "2tr1knot") %>%
  mutate(statistic = ifelse(p.fdr < 0.05 & statistic > 0, statistic, 0)) %>%
  build_overlay("statistic", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

