---
title: 'Comparing measures of pattern similarity using DMCC2 Stroop'
author: "mike freund"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    highlight: zenburn
---


# About

Here 12 different forms of similarity are considered.

* __3 measures of similarity__
    * linear **correlation**
    * **euclidean** distance
    * **standardized euclidean** distance

* __2 methods of estimation__
    * **vanilla** RSA: assessing similarity of patterns* *across scanning run**
    * **"cross-validated"** RSA: tweaking cross-run estimation procedure so that measures become *unbiased*

* __2 normalizing transforms__
    * **"raw"**, or un-normalized
    * spatially **prewhitened**, or "multivariate noise normalized"

For definition and discussion of these measures, see __./misc/measures_of_similarity.pdf__.

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = 'center',
  fig.width = 11.5, fig.fullwidth = TRUE
)

set.seed(0)

library(here)
library(magrittr)
library(dplyr)
library(tidyr)
library(data.table)
library(mikeutils)
library(lme4)
library(lmerTest)
library(ggplot2)
library(ggbeeswarm)
library(ggridges)
library(viridis)
library(grid)
library(gridExtra)
library(cowplot)
library(cifti)
library(gifti)
library(abind)
library(boot)

source(here("src", "setup.R"))


## settings ----

theme_set(theme_minimal(base_size = 14))


## data ----

## parcel-wise similarity matrices

## vanilla RSA

vanil <- 
  abind(
    abind(
      readRDS(here("out", "rsa", "observed", "rmatrix_vanilla_corr_shaefer400_baseline_Congruency_EVENTS_censored.rds")),
      readRDS(here("out", "rsa", "observed", "rmatrix_vanilla_eucl_shaefer400_baseline_Congruency_EVENTS_censored.rds")),
      readRDS(here("out", "rsa", "observed", "rmatrix_vanilla_neuc_shaefer400_baseline_Congruency_EVENTS_censored.rds")),
      rev.along = 0
    ),
    abind(
      readRDS(here("out", "rsa", "observed", "rmatrix_vanilla_corr_shaefer400_proactive_Congruency_EVENTS_censored.rds")),
      readRDS(here("out", "rsa", "observed", "rmatrix_vanilla_eucl_shaefer400_proactive_Congruency_EVENTS_censored.rds")),
      readRDS(here("out", "rsa", "observed", "rmatrix_vanilla_neuc_shaefer400_proactive_Congruency_EVENTS_censored.rds")),
      rev.along = 0
    ),
    rev.along = 0
  )

names(dimnames(vanil)) <- c(".row", ".col", "norma", "knot", "parcel", "subj", "measure", "session")
dimnames(vanil)$measure <- c("corr", "eucl", "neuc")
dimnames(vanil)$session <- c("baseline", "proactive")

cross <- 
  abind(
    abind(
      readRDS(here("out", "rsa", "observed", "rmatrix_crossva_corr_shaefer400_baseline_Congruency_EVENTS_censored.rds")),
      readRDS(here("out", "rsa", "observed", "rmatrix_crossva_eucl_shaefer400_baseline_Congruency_EVENTS_censored.rds")),
      readRDS(here("out", "rsa", "observed", "rmatrix_crossva_neuc_shaefer400_baseline_Congruency_EVENTS_censored.rds")),
      rev.along = 0
    ),
    abind(
        readRDS(here("out", "rsa", "observed", "rmatrix_crossva_corr_shaefer400_proactive_Congruency_EVENTS_censored.rds")),
        readRDS(here("out", "rsa", "observed", "rmatrix_crossva_eucl_shaefer400_proactive_Congruency_EVENTS_censored.rds")),
        readRDS(here("out", "rsa", "observed", "rmatrix_crossva_neuc_shaefer400_proactive_Congruency_EVENTS_censored.rds")),
        rev.along = 0
      ),
    rev.along = 0
  )

names(dimnames(cross)) <- c(".row", ".col", "norma", "knot", "parcel", "subj", "measure", "session")
dimnames(cross)$measure <- c("corr", "eucl", "neuc")
dimnames(cross)$session <- c("baseline", "proactive")

## filter subjs by those with data

has.stats <- apply(vanil, "subj", function(.) !any(is.na(c(.))))
subjs.with.stats <- names(has.stats)[has.stats]

## TODO
## these subjects do not have stats
# setdiff(c(subjs.analysis, subjs.development), subjs.with.stats)
## could be due to
##  - no 3d+t gifti -> have RAs run
##  - no beta gifti -> diagnose problem and fit

subjs.analysis <- intersect(subjs.analysis, subjs.with.stats)
subjs.development <- intersect(subjs.development, subjs.with.stats)
subjs.bad <- intersect(subjs.bad, subjs.with.stats)

vanil <- vanil[, , , , , subjs.analysis, , ]
cross <- cross[, , , , , subjs.analysis, , ]

## palettes ----

colors.norma <- c(raw = "grey30", prw = "firebrick")
# colors.glm <- c("2tr1knot" = "#d95f02", "1tr1knot" = "#1b9e77")
colors.celltype <- c(I_I = "#e41a1c", C_C = "#377eb8", I_C = "#4daf4a")

## transformations

vanil[, , , , , , "corr", ] <- atanh(vanil[, , , , , , "corr", ])  ## fisher transform
vanil[, , , , , , c("eucl", "neuc"), ] <- -vanil[, , , , , , c("eucl", "neuc"), ]  ## negate for interpretability
cross[, , , , , , c("eucl", "neuc"), ] <- -cross[, , , , , , c("eucl", "neuc"), ]
# cross.pro[, , , , , , c("eucl", "neuc")] <- -cross.pro[, , , , , , c("eucl", "neuc")]


## extract target knot ----

vanil <- vanil[, , , "knot2", , , , ]
cross <- cross[, , , "knot2", , , , ]

# cross <- cross[, , , c("knot3", "knot4"), , , ] %>% apply(c(".row", ".col", "norma", "parcel", "subj", "measure"), mean)
# vanil <- vanil[, , , c("knot3", "knot4"), , , ] %>% apply(c(".row", ".col", "norma", "parcel", "subj", "measure"), mean)


```


# 1. Impact of prewhitening

* baseline PC50 only

* define function $\mathit{cell}(\text{rowname}, \text{colname})$: takes row and column names as input, returns value of cell

* vanilla contrast: 

\[
\text{contrast}_\textit{vanilla} = 
\mathit{cell}(\text{PC50InConRun1}, \text{PC50InConRun2}) -
[\mathit{cell}(\text{PC50InConRun1}, \text{PC50ConRun2}) +
  \mathit{cell}(\text{PC50InConRun2}, \text{PC50ConRun1})]/2
\]

This is tested against zero (over subjects): $\text{contrast}_\textit{vanilla} > 0$

* cross-validated contrast:

\[
\text{contrast}_\textit{cross-validated} = \mathit{cell}(\text{PC50InCon}, \text{PC50Con})
\]

The invidiual cells themselves represent contrasts.

These can be tested against zero (over subjects): $\text{contrast}_\textit{cross-validated} > 0$



```{r prewhitening_wrangle}

## vanilla measures ----

inds.pc50.run1 <- grep("PC50.*_run1", conds.run)
inds.pc50.run2 <- grep("PC50.*_run2", conds.run)

vanil.pc50 <- 
  
  vanil[5:8, 1:4, , , , , "baseline"] %>%  ## btw run only

  reshape2::melt(value.name = "simil") %>% 
  
  group_by(norma, parcel, subj) %>%   ## standardize within subjects
  mutate(simil.sd = simil / sd(simil), method = "vanil")

## get only pc50 and create celltype col

vanil.pc50$.row <- gsub("_run.$", "", vanil.pc50$.row)
vanil.pc50$.col <- gsub("_run.$", "", vanil.pc50$.col)

vanil.pc50 <- vanil.pc50 %>% filter(grepl("PC50", .row), grepl("PC50", .col))

vanil.pc50$celltype <- ifelse(
  grepl("PC50InCon", vanil.pc50$.row) & grepl("PC50InCon", vanil.pc50$.col), "I_I",
  ifelse(
    grepl("PC50Con", vanil.pc50$.row) & grepl("PC50Con", vanil.pc50$.col), "C_C", "I_C"
  )
)

vanil.pc50 <- vanil.pc50 %>%
  
  group_by(subj, parcel, celltype, norma, measure) %>%
  summarize(simil = mean(simil), simil.sd = mean(simil.sd)) %>%  ## average within celltype
  
  group_by(subj, parcel, norma, measure) %>%  ## perform contrast
  summarize(
    simil = simil[celltype == "I_I"] - simil[celltype == "I_C"],
    simil.sd = simil.sd[celltype == "I_I"] - simil.sd[celltype == "I_C"]
    )

## crossvalidated measures ----

## standardize measures within subjects

## correlation
cross.corr <- cross[, , , , , "corr", "baseline"] %>% 
  reshape2::melt(value.name = "simil") %>% 
  mutate(method = "cross", measure = "corr")
cross.corr <- cross.corr %>% group_by(norma, parcel, subj) %>% mutate(simil.sd = simil / sd(simil))

## euclidean and normalized euclidean
cross.eucls <- -cross[, , , , , c("eucl", "neuc"), "baseline"] %>%  ## un-negate (larger distance, more different!)
  reshape2::melt(value.name = "simil") %>%
  mutate(method = "cross")
cross.eucls <- cross.eucls %>% group_by(norma, parcel, subj) %>% mutate(simil.sd = simil / sd(simil))

## get pc50 only and bind

cross.pc50 <- bind_rows(cross.corr, cross.eucls)
cross.pc50 <- cross.pc50[cross.pc50$.row == "PC50InCon" & cross.pc50$.col == "PC50Con", ]

## bind ----

rsa.pc50 <- bind_rows(
  vanil.pc50 %>% mutate(method = "vanil"),
  cross.pc50 %>% select(-.row, -.col)
)


## take out trash ----

rm(vanil.pc50, cross.corr, cross.eucls, cross.pc50)
gc()


## add factor cols ----

## bind and add factor cols

rsa.pc50$network <- ""  ## add networks
for (network.i in networks) rsa.pc50$network[grepl(network.i, rsa.pc50$parcel)] <- network.i

rsa.pc50$parcel.num <- match(rsa.pc50$parcel, parcellation$key)

```

## stats

```{r prewhitening_mean_contrasts}

rsa.pc50 %>%
  
  filter(parcel.num %in% dmcc34) %>%
  group_by(subj, norma, measure, method) %>%
  summarize(simil = mean(simil)) %>%
  
  ggplot(aes(norma, simil, fill = norma)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = subj), alpha = 0.2) +
  geom_boxplot(notch = TRUE, width = 0.25) +
  
  facet_wrap(vars(method, measure), scales = "free_y") +
  
  scale_fill_manual(values = colors.norma) +
  
  theme_light() +
  theme(legend.position = "none")

rsa.pc50 %>%
  
  filter(parcel.num %in% dmcc34) %>%
  group_by(subj, norma, measure, method) %>%
  summarize(simil.sd = mean(simil.sd)) %>%
  
  ggplot(aes(norma, simil.sd, fill = norma)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = subj), alpha = 0.2) +
  geom_boxplot(notch = TRUE, width = 0.25) +
  
  facet_wrap(vars(method, measure), scales = "free_y") +
  
  scale_fill_manual(values = colors.norma) +
  
  theme_light() +
  theme(legend.position = "none")
  
# rsa.pc50 %>%
#   
#   filter(parcel.num %in% dmcc34) %>%
#   group_by(norma, measure, method, subj) %>%
#   summarize(simil = mean(simil)) %>%
#   mutate(simil = simil / sd(simil)) %>%
#   
#   ggplot(aes(norma, simil, fill = norma)) +
#   geom_hline(yintercept = 0) +
#   geom_line(aes(group = subj), alpha = 0.2) +
#   geom_boxplot(notch = TRUE, width = 0.25) +
#   
#   facet_wrap(vars(method, measure), scales = "free_y") +
#   
#   scale_fill_manual(values = colors.norma) +
#   
#   theme_light() +
#   theme(legend.position = "none")


rsa.pc50 %>%
  
  group_by(parcel, parcel.num, norma, measure, method, subj) %>%
  summarize(simil = mean(simil.sd)) %>%
  
  summarize(
    ssr = wilcox.test(simil, alternative = "greater")$statistic,
    p.value = wilcox.test(simil, alternative = "greater")$p.value
    ) %>%
  group_by(parcel, parcel.num, measure, method) %>%
  mutate(is.sig = any(p.value < 0.05), p.value = NULL) %>%
  
  pivot_wider(names_from = "norma", values_from = "ssr") %>%
  
  ggplot(aes(raw, prw, fill = parcel.num %in% dmcc34)) +
  geom_abline() +
  geom_hline(yintercept = 2) +
  geom_vline(xintercept = 2) +
  # geom_point(shape = 21, color = "white", size = 2) +
  geom_point(aes(alpha = is.sig), shape = 21, color = "white", size = 2) +
  
  facet_grid(vars(method), vars(measure)) +
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.2)) +
  
  theme_light() +
  theme(legend.position = "none")


```



```{r prewhitening_brains}

onesamp <- rsa.pc50 %>%
  group_by(parcel, norma, measure, method, network, parcel.num) %>%
  summarize(
    m = mean(simil.sd),
    ssr = wilcox.test(simil.sd, alternative = "greater")$statistic,
    p.value = wilcox.test(simil.sd, alternative = "greater")$p.value
    ) %>%
  group_by(norma, measure, method, network) %>%
  mutate(p.fdr = p.adjust(p.value, method = "fdr"))

paired <- rsa.pc50 %>%
  group_by(parcel, norma, measure, method, network, parcel.num) %>%
  mutate(simil.sd = simil.sd / sd(simil.sd)) %>%
  select(-simil) %>%
  pivot_wider(names_from = "norma", values_from = "simil.sd") %>%
  summarize(
    m = mean(prw) - mean(raw),
    ssr = wilcox.test(prw, raw, paired = TRUE)$statistic,
    p.value = wilcox.test(prw, raw, paired = TRUE)$p.value,
    ) %>%
  group_by(measure, method, network) %>%
  mutate(p.fdr = p.adjust(p.value, method = "fdr"))

# table(interaction(a$norma, a$measure, a$method), a$p.fdr < 0.05)

## add cols for plotting

onesamp %<>%
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
    )

paired %<>%
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
    )

```


## brains

* the t statistics from above are plotted on brains here.

* thresholded at FDR-corrected whole-brain $\alpha = 0.05$

### one-sample wilcox stats

#### vanilla measures

##### correlation, raw

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "vanil", measure == "corr", norma == "raw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### correlation, prewhitened

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "vanil", measure == "corr", norma == "prw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### euclidean, raw

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "vanil", measure == "eucl", norma == "raw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### euclidean, prewhitened

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "vanil", measure == "eucl", norma == "prw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```


##### standardized euclidean, raw

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "vanil", measure == "neuc", norma == "raw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### standardized euclidean, prewhitened

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "vanil", measure == "neuc", norma == "prw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

#### cross-validated measures

##### correlation, raw

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "cross", measure == "corr", norma == "raw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### correlation, prewhitened

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "cross", measure == "corr", norma == "prw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### euclidean, raw

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "cross", measure == "eucl", norma == "raw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### euclidean, prewhitened

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "cross", measure == "eucl", norma == "prw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### standardized euclidean, raw

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "cross", measure == "neuc", norma == "raw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### standardized euclidean, prewhitened

```{r, fig.height = 7, fig.width = 11}

onesamp %>%
  filter(method == "cross", measure == "neuc", norma == "prw") %>%
  mutate(ssr = ifelse(p.fdr < 0.05, ssr, 0)) %>%
  build_overlay("ssr", template = schaefer) %>%
  plot_surface(underlay = hcp)

```


### two-sample, paired wilcox stats

* plotted: standardized mean difference between prewhitened and raw measures (prw - raw), within subject

* thresholded at uncorrected p = 0.05

#### vanilla measures

##### correlation

```{r, fig.height = 7, fig.width = 11}

paired %>%
  filter(method == "vanil", measure == "corr") %>%
  mutate(m = ifelse(p.value < 0.05, m, 0)) %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### euclidean

```{r, fig.height = 7, fig.width = 11}

paired %>%
  filter(method == "vanil", measure == "eucl") %>%
  mutate(m = ifelse(p.value < 0.05, m, 0)) %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### standardized euclidean

```{r, fig.height = 7, fig.width = 11}

paired %>%
  filter(method == "vanil", measure == "neuc") %>%
  mutate(m = ifelse(p.value < 0.05, m, 0)) %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

#### cross-validated measures

##### correlation

```{r, fig.height = 7, fig.width = 11}

paired %>%
  filter(method == "cross", measure == "corr") %>%
  mutate(m = ifelse(p.value < 0.05, m, 0)) %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### euclidean

```{r, fig.height = 7, fig.width = 11}

paired %>%
  filter(method == "cross", measure == "eucl") %>%
  mutate(m = ifelse(p.value < 0.05, m, 0)) %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### standardized euclidean

```{r, fig.height = 7, fig.width = 11}

paired %>%
  filter(method == "cross", measure == "neuc") %>%
  mutate(m = ifelse(p.value < 0.05, m, 0)) %>%
  build_overlay("m", template = schaefer) %>%
  plot_surface(underlay = hcp)

```


# 2. Impact of cross-validation on sensitivity to number of trials

### Mean (whole-brain) similarity matrices

_cf._, the 'vanilla' matrices displayed in the __./rois_rsa_vanilla/.__ analyses

```{r crossvalidated_matrices, fig.height = 10}

allsim <- apply(cross, c(".row", ".col", "norma", "measure", "session"), mean)


grid.arrange(
  
  allsim[, , "raw", "corr", "baseline"] %>% plot_matrix + labs(title = "correlation, bas"),
  allsim[, , "raw", "corr", "proactive"] %>% plot_matrix + labs(title = "correlation, pro"),
  
  allsim[, , "raw", "eucl", "baseline"] %>% plot_matrix + labs(title = "euclidean, bas"),
  allsim[, , "raw", "eucl", "proactive"] %>% plot_matrix + labs(title = "euclidean, pro"),
  
  allsim[, , "raw", "neuc", "baseline"] %>% plot_matrix + labs(title = "standardized euclidean, bas"),
  allsim[, , "raw", "neuc", "proactive"] %>% plot_matrix + labs(title = "standardized euclidean, pro"),
  
  ncol = 2
  
)

```


**a rough analysis of sensitivity of measures to trial balancing**

* focus on proactive data: not much going on in terms of observed stroop effects

* BiasInCon in proactive has 54 trials per run

* all other trial types (PC50Con, PC50InCon, biasCon), in proactive have 18 trials per run

* If a measure is sensitive to trial counts, BiasInCon should correlate more strongly with all other patterns, regardless of pc / congruency.

* Get average/sum of similarities of biasInCon and PC50InCon to congruent patterns:

\[\text{biasInCon}\sim\text{allCon} = \mathit{simil}(\text{BiasInCon}, \text{PC50Con}) + \mathit{simil}(\text{BiasInCon}, \text{biasCon})\]
\[\text{PC50InCon}\sim\text{allCon} = \mathit{simil}(\text{PC50InCon}, \text{PC50Con}) + \mathit{simil}(\text{PC50InCon}, \text{biasCon})\]

    * these indicate the average *cross-congruency* similarity for biasInCon and PC50InCon
    
* contrast: $(\text{biasInCon}\sim\text{allCon}) - (\text{PC50InCon}\sim\text{allCon})$?
    * will be positive if measure sensitive to number of trials.


```{r crossvalidated_stats}

## wrangle ----

## vanilla measures

vanil.d <- vanil[1:4, 5:8, , , , , ] %>% reshape2::melt(value.name = "simil") %>% mutate(method = "vanil")

vanil.d$.row <- gsub("_run.$", "", vanil.d$.row)
vanil.d$.col <- gsub("_run.$", "", vanil.d$.col)

vanil.d$pc <- ifelse(
  grepl("PC50", vanil.d$.row) & grepl("PC50", vanil.d$.col), "pc50_pc50",
  ifelse(
    grepl("bias", vanil.d$.row) & grepl("bias", vanil.d$.col), "bias_bias",
    "pc50_bias"
  )
)
vanil.d$congruency <- ifelse(
  grepl("InCon", vanil.d$.row) & grepl("InCon", vanil.d$.col), "I_I",
  ifelse(
    grepl("biasCon|PC50Con", vanil.d$.row) & grepl("biasCon|PC50Con", vanil.d$.col), "C_C",
    "I_C"
  )
)

vanil.d <- vanil.d %>% filter(congruency == "I_C")

vanil.d$condition <- ifelse(
    grepl("biasInCon", vanil.d$.row) | grepl("biasInCon", vanil.d$.col), "biasI_C",
    ifelse(
      grepl("PC50InCon", vanil.d$.row) | grepl("PC50InCon", vanil.d$.col), "pc50I_C", NA
    )
)

# sum(is.na(vanil.d$condition))

vanil.d <- vanil.d %>%
  group_by(parcel, subj, measure, session, method, condition, norma) %>%
  summarize(simil = mean(simil))  ## summarize by condiiton


## cross-validated measures

cross.d <- cross %>% reshape2::melt(value.name = "simil") %>% mutate(method = "cross")

cross.d$pc <- ifelse(
  grepl("PC50", cross.d$.row) & grepl("PC50", cross.d$.col), "pc50_pc50",
  ifelse(
    grepl("bias", cross.d$.row) & grepl("bias", cross.d$.col), "bias_bias",
    "pc50_bias"
  )
)
cross.d$congruency <- ifelse(
  grepl("InCon", cross.d$.row) & grepl("InCon", cross.d$.col), "I_I",
  ifelse(
    grepl("biasCon|PC50Con", cross.d$.row) & grepl("biasCon|PC50Con", cross.d$.col), "C_C",
    "I_C"
  )
)

cross.d <- cross.d %>% filter(congruency == "I_C")

cross.d$condition <- ifelse(
    grepl("biasInCon", cross.d$.row) | grepl("biasInCon", cross.d$.col), "biasI_C",
    ifelse(
      grepl("PC50InCon", cross.d$.row) | grepl("PC50InCon", cross.d$.col), "pc50I_C", NA
    )
)

# sum(is.na(cross.d$condition))

cross.d <- cross.d %>% 
  group_by(parcel, subj, measure, session, method, condition, norma) %>%
  summarize(simil = mean(simil))  ## summarize by condiiton

## bind

d <- bind_rows(cross.d, vanil.d)

## stats

d.stats <- d %>%
  
  .[.$session == 'proactive' & .$norma == "raw", ] %>%
  
  pivot_wider(names_from = "condition", values_from = "simil") %>%
  group_by(parcel, measure, method) %>%
  summarize(
    statistic = t.test(biasI_C, pc50I_C, paired = TRUE)$statistic,
    p.value = t.test(biasI_C, pc50I_C, paired = TRUE)$p.value
    ) %>%
  group_by(measure, method) %>%
  mutate(p.fdr = p.adjust(p.value, method = "fdr"))


d.stats <- d.stats %>% 
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
    )

```



```{r crossvalidated_plots}

d %>%
  
  .[.$session == 'proactive' & .$norma == "raw", ] %>%
  
  group_by(subj, measure, session, condition, method) %>%
  summarize(simil = mean(simil)) %>%
  pivot_wider(names_from = "condition", values_from = "simil") %>%
  
  ggplot(aes(biasI_C, pc50I_C)) +
  geom_abline() +
  geom_point(shape = 21, fill = "black", color = "grey50", size = 2) +
  
  facet_wrap(vars(method, measure), scales = "free") +
  labs(
    title = "cross-congruency similarities per subject (aggregated across all parcels, proactive)",
    x = "biasInCon~allCon",
    y = "PC50InCon~allCon",
    caption = "line is unity"
    )


# d.stats %>%
#   
#   ggplot(aes(method, statistic)) +
#   geom_boxplot(aes(fill = measure), position = position_dodge())

d.stats %>%
  
  select(-p.value, -p.fdr) %>%
  pivot_wider(names_from = "method", values_from = "statistic") %>%
  
  ggplot(aes(cross, vanil)) +
  geom_abline() +
  geom_point(aes(color = match(parcel, parcellation$key) %in% dmcc34)) +
  geom_point(
    data = . %>% group_by(measure) %>% summarize(cross = mean(cross), vanil = mean(vanil)),
    size = 5 
  ) +
  
  facet_wrap(vars(measure), scales = "free") +
  
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey50"), name = "in dmcc34") +

  labs(
    title = "cross-congruency similarity CONTRAST per parcel (paired sample t-test over subjects, proactive)",
    x = "cross-validated methods",
    y = "vanilla methods",
    caption = "line is unity\n black dot is centroid"
    )

```

Cross-congruency similarity contrast, table of parcel counts with significant effect ("TRUE"; uncorrected):

```{r echo = TRUE}

with(d.stats, table(interaction(measure, method), p.value < 0.05))

# parcels.eucl <- d.stats %>% filter(p.value < 0.05, measure == "eucl", method == "cross") %>% pull(parcel)
# parcels.corr <- d.stats %>% filter(p.value < 0.05, measure == "corr", method == "cross") %>% pull(parcel)
# parcels.neuc <- d.stats %>% filter(p.value < 0.05, measure == "neuc", method == "cross") %>% pull(parcel)
# 
# setdiff(parcels.eucl, parcels.corr)
# intersect(parcels.neuc, parcels.corr)
# setdiff(parcels.neuc, parcels.corr)
# setdiff(parcels.corr, parcels.neuc)

```

##### correlation, cross-validated

```{r, fig.height = 7, fig.width = 11}

d.stats %>%
  
  filter(method == "cross", measure == "corr") %>%
  mutate(t.value = ifelse(p.value < 0.05, statistic, 0)) %>%
  
  build_overlay("t.value", template = schaefer) %>%
  plot_surface(underlay = hcp)

```


##### euclidean, cross-validated

```{r, fig.height = 7, fig.width = 11}

d.stats %>%
  
  filter(method == "cross", measure == "eucl") %>%
  mutate(t.value = ifelse(p.value < 0.05, statistic, 0)) %>%
  
  build_overlay("t.value", template = schaefer) %>%
  plot_surface(underlay = hcp)

```

##### standardized euclidean, cross-validated

```{r, fig.height = 7, fig.width = 11}

d.stats %>%
  
  filter(method == "cross", measure == "neuc") %>%
  mutate(t.value = ifelse(p.value < 0.05, statistic, 0)) %>%
  
  build_overlay("t.value", template = schaefer) %>%
  plot_surface(underlay = hcp)

```


# 3. Extended analyses of prewhitening on cross-validated measures

## (a) between-subject correlations with alternative measures

### stroop effects (I--C discriminability) across item types (pc50, bias)

```{r}

n.resamp <- 1000

cross.d <- cross.d %>%
  ungroup %>%
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
  )

cross.d2 <- cross.d %>% 
  
  .[.$num.roi %in% dmcc34 & .$session == "baseline", ] %>%
  
  pivot_wider(names_from = c("condition", "norma"), values_from = "simil")
  
# df <- cross.d2 %>% filter(parcel == "LH_Vis_22", measure == "corr")  

```


```{r}

g <- with(cross.d2, paste0(measure, ".", parcel))
l <- split(cross.d2, g)

boot.cross.cell <- l %>% 
  lapply(
    function(.) {
      boot(
        ., 
        statistic = function(dat, ind) {
          dat.i <- dat[ind, ]
          r.raw <- cor(dat.i$biasI_C_raw, dat.i$pc50I_C_raw)
          r.prw <- cor(dat.i$biasI_C_prw, dat.i$pc50I_C_prw)
          r.prw^2 - r.raw^2
          },
        R = n.resamp
      )
    }
  )

res.cross.cell <- boot.cross.cell %>% vapply(function(.) boot.ci(., type = "bca")$bca[4:5], numeric(2))
res.cross.cell <- rbind(res.cross.cell, boot.cross.cell %>% vapply(function(.) .$t0, numeric(1)))
res.cross.cell <- as.data.frame(t(res.cross.cell))
res.cross.cell %<>% cbind(reshape2::colsplit(rownames(.), "\\.", c("measure", "parcel")))
res.cross.cell %<>% rename(lb = V1, ub = V2, m = V3)


res.cross.cell %>%
  
  ggplot(aes(parcel, m)) +
  
  geom_hline(yintercept = 0) +
  geom_point() +
  geom_errorbar(aes(ymin = lb, ymax = ub)) +
  
  facet_grid(vars(measure)) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```


### association with RT



```{r}

behav.data <- pretrial %>% filter(session == "bas", subj %in% subjs.analysis, rt > 500, rt < 2500, acc == 1, !is.na(rt))
fit <- lmer(rt ~ trial.type + (trial.type | subj), behav.data)
stroop.effs <- data.frame(subj = rownames(coef(fit)$subj), stroop.rt = coef(fit)$subj$trial.typei)

vanil.d2 <- vanil.d %>% 
  
  ungroup %>%
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
  ) %>%
  
  .[.$num.roi %in% dmcc34 & .$session == "baseline", ] %>%
  
  pivot_wider(names_from = c("condition", "norma"), values_from = "simil")

d2 <- bind_rows(vanil.d2, cross.d2)

d2 <- inner_join(d2, stroop.effs, by = "subj")

d2$I_C_raw <- (d2$biasI_C_raw + d2$pc50I_C_raw) / 2
d2$I_C_prw <- (d2$biasI_C_prw + d2$pc50I_C_prw) / 2

g <- with(d2, paste0(method, "." ,measure, ".", parcel))
l <- split(d2, g)

boot.stroop <- l %>% 
  lapply(
    function(.) {
      boot(
        ., 
        statistic = function(dat, ind) {
          dat.i <- dat[ind, ]
          r.raw <- cor(dat.i$I_C_raw, dat.i$stroop.rt)
          r.prw <- cor(dat.i$I_C_prw, dat.i$stroop.rt)
          r.prw^2 - r.raw^2
          },
        R = n.resamp
      )
    }
  )

res.stroop <- boot.stroop %>% vapply(function(.) boot.ci(., type = "bca")$bca[4:5], numeric(2))
res.stroop <- rbind(res.stroop, boot.stroop %>% vapply(function(.) .$t0, numeric(1)))
res.stroop <- as.data.frame(t(res.stroop))
res.stroop %<>% cbind(reshape2::colsplit(rownames(.), "\\.", c("method", "measure", "parcel")))
res.stroop %<>% rename(lb = V1, ub = V2, m = V3)

res.stroop %>%
  
  ggplot(aes(parcel, m)) +
  
  geom_hline(yintercept = 0) +
  geom_point() +
  geom_errorbar(aes(ymin = lb, ymax = ub)) +
  
  facet_grid(cols = vars(method, measure)) +

  # theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  
  coord_flip()


```


## (b) noise ceilings



```{r}

## crossvalidated measures ----

## standardize measures within subjects

## euclidean and normalized euclidean

cross.eucls <- cross[, , , , , c("eucl", "neuc"), ]

for (row.i in seq_len(nrow(cross.eucls))) {
  for (col.j in seq_len(ncol(cross.eucls))) {
    if (row.i <= col.j) cross.eucls[row.i, col.j, , , , , ] <- NA
  }
}

cross.eucls <- cross.eucls %>%
  reshape2::melt(value.name = "simil", na.rm = TRUE) %>%
  mutate(method = "cross")
cross.eucls <- cross.eucls %>% group_by(norma, parcel, subj) %>% mutate(simil.sd = simil / sd(simil))

## add factor cols ----

## bind and add factor cols

cross.eucls$network <- ""  ## add networks
for (network.i in networks) cross.eucls$network[grepl(network.i, cross.eucls$parcel)] <- network.i

cross.eucls$parcel.num <- match(cross.eucls$parcel, parcellation$key)

cross.eucls <- cross.eucls %>%   
  ungroup %>%
  mutate(
    hemi = substr(parcel, 1, 1),
    num.roi = match(parcel, parcellation$key)
  )

noise.ceilings <- cross.eucls %>%
  
  group_by(.row, .col, norma, parcel, measure, method, session) %>%
  mutate(best.model = mean(simil.sd)) %>%
  
  group_by(norma, parcel, subj, measure, network, session) %>%
  summarize(ub = cor(best.model, simil.sd))


noise.ceilings %>%
  
  group_by(parcel, measure, network, norma, session) %>%
  summarize(ub = mean(ub)) %>%
  
  pivot_wider(names_from = "norma", values_from = "ub") %>%
  
  ggplot(aes(raw, prw)) +
  geom_abline() +
  geom_point(aes(color = match(parcel, parcellation$key) %in% dmcc34)) +
  
  geom_point(
    data = . %>% group_by(measure) %>% summarize(prw = mean(prw), raw = mean(raw)),
    size = 5 
  ) +
  geom_point(
    data = . 
      %>% filter(match(parcel, parcellation$key) %in% dmcc34)
      %>% group_by(measure) %>% summarize(prw = mean(prw), raw = mean(raw)),
    size = 5,
    color = "red"
  ) +

  facet_grid(vars(session), vars(measure)) +
  
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey50"), name = "in dmcc34")

```

